---
title: "Introduction to AzureRMR"
author: Hong Ooi
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to AzureRMR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{utf8}
---

AzureRMR is a package for interacting with Azure Resource Manager: authenticate, list subscriptions, manage resource groups, deploy and delete templates and resources. It calls the Resource Manager [REST API](https://docs.microsoft.com/en-us/rest/api/resources) directly, so you don't need to have PowerShell or Python installed.

As a general-purpose interface to Azure Resource Manager (ARM), you can use AzureRMR to work with nearly any Azure object that ARM can handle: subscriptions, resource groups, resources, templates and so on. The things you can do include:

- Create a new resource
- Carry out arbitrary operations on a resource
- Delete a resource
- Deploy a template
- Delete a template, and, optionally, any resources that it created
- Create and delete resource groups (if you gave your service principal subscription-level access)


## Authentication

AzureRMR uses a similar authentication process to the [Azure CLI](https://docs.microsoft.com/en-us/cli/azure/?view=azure-cli-latest). The first time you authenticate with a given Azure Active Directory tenant, you call `create_az_login("tenant_name")`. R will display a code and prompt you to visit the Microsoft login page in your browser. You then enter the code along with your Active Directory credentials, which completes the authentication process. The returned Resource Manager client object is also saved on your machine, and can be retrieved in subsequent R sessions with `get_az_login("tenant_name")`. AzureRMR will automatically handle details like refreshing your credentials.

This achieves two things: first, you only have to manually authenticate once; and second, it saves you from having to create a service principal (although you can also do that if so desired).


## Logging in

Let's look at a sample workflow involving AzureRMR. The first thing to do is authenticate with Azure, which we do with `get_az_login` as described above. This in fact uses the same app ID as the Azure CLI, so it should work anywhere you can use the CLI.

```{r, eval=FALSE}
library(AzureRMR)

# if this is the first time you're logging in to this tenant
az <- create_az_login("myaadtenant.onmicrosoft.com")

# you can also supply just the tenant name or ID
az <- create_az_login("myaadtenant")

# for subsequent sessions
az <- get_az_login("myaadtenant")

# you can also list the tenants that you've previously authenticated with
list_az_logins()
```

As a convenience feature, you can also call `get_az_login("tenant_name")` on first login rather than `create_az_login`. In this case, AzureRMR will detect that you don't have a saved client object and create it for you.


## Subscriptions and resource groups

AzureRMR allows you to work with your subscriptions and resource groups. Note that if you created your service principal via the cloud shell, as described in this vignette, you probably only have access to one subscription. Regardless, you can list all subscriptions that you can work with:

```{r, eval=FALSE}
# all subscriptions
az$list_subscriptions()
#$`5710aa44-281f-49fe-bfa6-69e66bb55b11`
#<Azure subscription 5710aa44-281f-49fe-bfa6-69e66bb55b11>
#  authorization_source: Legacy
#  name: Visual Studio Ultimate with MSDN
#  policies: list(locationPlacementId, quotaId, spendingLimit)
#  state: Enabled
#---
#  Methods:
#    create_resource_group, delete_resource_group, get_provider_api_version, get_resource_group,
#    list_locations, list_resource_groups, list_resources
#
#$`e26f4a80-370f-4a77-88df-5a8d291cd2f9`
#<Azure subscription e26f4a80-370f-4a77-88df-5a8d291cd2f9>
#  authorization_source: RoleBased
#  name: ADLTrainingMS
#  policies: list(locationPlacementId, quotaId, spendingLimit)
#  state: Enabled
#---
#  Methods:
#    create_resource_group, delete_resource_group, get_provider_api_version, get_resource_group,
#    list_locations, list_resource_groups, list_resources
#
#...
```

Notice that AzureRMR is based on R6 classes, where methods are part of the object itself (much like objects in C++, C# and Java). Thus `list_subscriptions` is a member of the `az` object, and we call it with `az$list_subscriptions()`. R6 is used because it allows objects to have persistent state; in this case, the objects in R represent corresponding objects in Azure.

The `list_subscriptions()` call returns a list of subscription objects. You can retrieve the details for a single subscription with `get_subscription`:

```{r, eval=FALSE}
# get a subscription
(sub1 <- az$get_subscription("5710aa44-281f-49fe-bfa6-69e66bb55b11"))
#<Azure subscription 5710aa44-281f-49fe-bfa6-69e66bb55b11>
#  authorization_source: Legacy
#  name: Visual Studio Ultimate with MSDN
#  policies: list(locationPlacementId, quotaId, spendingLimit)
#  state: Enabled
#---
#  Methods:
#    create_resource_group, delete_resource_group, get_provider_api_version, get_resource_group,
#    list_locations, list_resource_groups, list_resources
```

A subscription object in turn has methods to get, create and delete resource groups (and also list all resource groups):

```{r, eval=FALSE}
(rg <- sub1$get_resource_group("rdev1"))
#<Azure resource group rdev1>
#  id: /subscriptions/5710aa44-281f-49fe-bfa6-69e66bb55b11/resourceGroups/rdev1
#  location: australiaeast
#  managed_by: NULL
#  properties: list(provisioningState)
#  tags: NULL
#---
#  Methods:
#    check, create_resource, delete, delete_resource, delete_template, deploy_template, get_resource,
#    get_template, list_resources, list_templates

# create and delete a resource group
test <- sub1$create_resource_group("test_group")
test$delete(confirm=FALSE)
```

## Resources and templates

Methods for working with resources and templates are exposed as part of the `az_resource_group` class. You can retrieve an existing resource/template, create a new one, or delete an existing one.

```{r, eval=FALSE}
(stor <- rg$get_resource(type="Microsoft.Storage/storageServices", name="rdevstor1"))
#<Azure resource Microsoft.Storage/storageAccounts/rdevstor1>
#  id: /subscriptions/5710aa44-281f-49fe-bfa6-69e66bb55b11/resourceGroups/rdev1/providers/Microsoft.Sto ...
#  identity: NULL
#  is_synced: TRUE
#  kind: Storage
#  location: australiasoutheast
#  managed_by: NULL
#  plan: NULL
#  properties: list(networkAcls, trustedDirectories, supportsHttpsTrafficOnly, encryption,
#    provisioningState, creationTime, primaryEndpoints, primaryLocation, statusOfPrimary)
#  sku: list(name, tier)
#  tags: list()
#---
#  Methods:
#    check, delete, do_operation, set_api_version, sync_fields, update
```

One benefit of the syntax that AzureRMR uses is that _method chaining_ works. This is the OOP version of pipelines, which most R users will recognise from the tidyverse.

```{r, eval=FALSE}
# use method chaining to get a resource without creating a bunch of intermediaries
# same result as above
stor <- az$
    get_subscription("5710aa44-281f-49fe-bfa6-69e66bb55b11")$
    get_resource_group("rdev1")$
    get_resource(type="Microsoft.Storage/storageServices", name="rdevstor1")
```

Once we have a resource, we can do _things_ with it, via the `do_operation()` method. In this case, we have a storage account. One of the things we can do with a storage account is retrieve its access keys:

```{r, eval=FALSE}
stor$do_operation("listKeys", http_verb="POST")
# $`keys`
# $`keys`[[1]]
# $`keys`[[1]]$`keyName`
# [1] "key1"
# 
# $`keys`[[1]]$value
# [1] "k0gGFi8LirKcDNe73fzwDzhZ2+4oRKzvz+6+Pfn2ZCKO/JLnpyBSpVO7btLxBXQj+j8MZatDTGZ2NXUItye/vA=="
# 
# $`keys`[[1]]$permissions
# [1] "FULL"
#...
```

Here is another example. If we have a virtual machine, we can start it, execute shell commands, and then shut it down again:

```{r, eval=FALSE}
vm <- rg$get_resource(type="Microsoft.Compute/virtualMachines",
    name="myVirtualMachine")

vm$do_operation("start", http_verb="POST") # may take a while
vm$do_operation("runCommand",
    body=list(
        commandId="RunShellScript", # RunPowerShellScript for Windows
        script=as.list("ifconfig > /tmp/ifconfig.out")
    ),
    encode="json",
    http_verb="POST")
vm$do_operation("powerOff", http_verb="POST")
```

For the types of operations you can carry out on a resource, consult the [Azure REST API documentation](https://docs.microsoft.com/en-us/rest/api/?view=Azure).

You can also interrogate the fields of a resource object; in particular the `properties` field can contain arbitrary information about an Azure resource. For example, a storage account's properties includes the endpoint URIs, and a virtual machine's properties includes its admin login details.

```{r, eval=FALSE}
# file and blob storage endpoint
stor$properties$primaryEndpoints$file
stor$properties$primaryEndpoints$blob

# OS profile for a VM: includes login details
vm$properties$osProfile
```


In a manner similar to resources, deploying a [template](https://www.red-gate.com/simple-talk/cloud/infrastructure-as-a-service/azure-resource-manager-arm-templates/) is just a matter of calling the resource group object's `deploy_template` method. This takes two arguments, `template` and `parameters`. Both arguments should be in JSON format: either the name of a JSON file, a character vector containing the JSON data, or a list containing the parsed JSON (via `jsonlite::toJSON`).

```{r, eval=FALSE}
vm_tpl <- rg$deploy_template("myNewVirtualMachine",
    template="vm_template.json",
    parameters=list(
        os="Linux",
        size="Standard_DS2_v2",
        username="ruser",
        publickey=readLines("~/id_rsa.pub")
    ))
```

Normally, deleting a template doesn't touch the resources it creates: it only deletes the template itself. However, AzureRMR optionally allows you to free any resources created when you delete a template. This is useful when managing complex objects like VMs, which actually consist of multiple individual resources in Azure (storage account, disk, network interface, etc). When you are done with the VM, deleting the template lets you free all these resources with a single command.

```{r, eval=FALSE}
vm_tpl$delete(free_resources=TRUE)
```


## Conclusion

This has been a quick introduction to the features of AzureRMR. Remember that this package is only meant to be a generic mechanism for working with Resource Manager. You can extend it to provide support for service-specific features; examples of packages that do this include [AzureVM](https://github.com/cloudyr/AzureVM) for [virtual machines](https://azure.microsoft.com/en-us/services/virtual-machines/), and [AzureStor](https://github.com/cloudyr/AzureStor) for [storage accounts](https://azure.microsoft.com/en-us/services/storage/). For more information, see the "Extending AzureRMR" vignette.


